\section{Introduction}
\label{sec:introduction}

Speech emotion recognition (SER) underpins affect-aware interfaces in education, pediatric healthcare, and human--robot interaction.
For \textit{children's} speech, the problem is compounded by wider fundamental-frequency ranges, immature articulation, and limited availability of spontaneous, labeled corpora.
Most child SER systems therefore rely on acted datasets recorded under studio conditions, where emotions are exaggerated and acoustically stereotyped.

This practice creates a silent validity crisis.
A model that achieves high accuracy on acted child speech may learn dataset-specific prototypes rather than generalizable emotional cues.
When such models encounter spontaneous child vocalizations---as in the FAU Aibo child--robot corpus---performance can collapse catastrophically.
Yet the field continues to report in-domain acted accuracy as the primary success metric.

\subsection{Motivation: Distribution Shift as the Central Variable}

Recent self-supervised learning (SSL) backbones such as WavLM~\cite{chen2022wavlm} provide powerful frame-level representations, but downstream design choices (pooling, adaptation, regularization) still determine whether emotion-relevant information is extracted or whether the model overfits to corpus artifacts.
Mean pooling discards temporal structure; unconstrained self-attention pooling can attend to arbitrary feature dimensions without physical interpretability.

Children's emotional expression is strongly linked to \textit{prosody}: pitch excursions, energy bursts, and rhythm irregularities that differ systematically from adult speech~\cite{fan2024children}.
We hypothesize that explicitly conditioning temporal attention on frame-level F0 and RMS energy provides an inductive bias that (i)~accelerates learning, (ii)~regularizes models on noisy spontaneous speech, and (iii)~anchors decisions to measurable acoustic events.

\subsection{Contributions}

This paper makes four contributions:

\begin{enumerate}
\item \textbf{A distribution-driven evaluation protocol} spanning acted child (C-BESD), spontaneous child (FAU Aibo), and adult (IEMOCAP) corpora, with speaker-disjoint splits and dual metrics (WA and UAR).
\item \textbf{Prosody-guided temporal importance pooling} fused with frozen WavLM and learnable 12-layer weighting, compared against a strictly parameter-matched self-attention baseline (111,105 parameters each).
\item \textbf{Quantification of the acted--spontaneous chasm}: zero-shot C-BESD$\rightarrow$FAU transfer yields only 19.56\% WA (vs.\ 92.78\% in-domain), with monotonic FD--accuracy degradation across age, style, and language shifts.
\item \textbf{An accuracy--explainability trade-off analysis}: prosody-guided pooling achieves APC$_{\text{wav}}{=}0.718$ with attention locked to energy peaks, justifying a 1.48pp WA sacrifice on acted data for deployable interpretability in safety-critical applications.
\end{enumerate}

\subsection{Paper Organization}

Section~\ref{sec:methodology} describes the model architecture and training protocol.
Section~\ref{sec:experiments_results} presents the full experimental matrix.
Section~\ref{sec:analysis_discussion} interprets results through three pillars: distribution-shift failure, prosody as structural regularization, and explainability.
Section~\ref{sec:conclusion} summarizes implications and future work.
